Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-03-31 ms.tex
bdm

Understanding why most SMBHs appear to be inactive requires
characterizing their gaseous environments.  Gas near the sphere of
influence of the SMBH, hereafter the `circumnuclear medium' (CNM),
controls the mass accretion rate, $\dot{M}$.  The accretion rate in
turn sets the luminosity of the black hole and the impact of this
energy and momentum output on larger physical scales (`feedback').
The presence of dense gas in the nuclear region may also lead to runaway
cooling, resulting in bursty episodes of star formation and AGN
activity (e.g.~\citealt{Ciotti&Ostriker07}; \citealt{Ciotti+10}).

Knowledge of how $\dot{M}$ depends on the SMBH mass, $\Mbh$, and other
properties of the galactic nucleus informs key questions related to
the co-evolution of SMBHs and their host galaxies with cosmic time
(e.g.~\citealt{Kormendy&Ho13}; \citealt{HeckmanBest:2014a}).  SMBH growth
in the low redshift universe is dominated by low mass black holes,
$M_{\bullet} \lesssim 10^{8}M_{\odot}$ (e.g.~\citealt{Heckman+04}), a
fact which is often attributed to the general trend of `cosmic
down-sizing' resulting from hierarchical structure growth (e.g,
\citealt{Gallo+08}).  However, the physical processes involved in
growing low mass blacks may be distinct from those operating at higher
SMBH masses and accretion rates in quasars.  A key open question is
whether SMBHs grow primarily by accreting gas fed in directly from
galactic or extra-galactic scales, or whether significant growth
results also from internal, secular processes such as local stellar
mass loss in the nuclear region.

A better understanding of what mechanisms regulate accretion onto
quiescent SMBHs would have implications for a variety of observations,
such as constraints on the occupation fraction of SMBHs in low mass
galaxies.  \citet{Miller+15} use the average relationship between the
nuclear X-ray luminosities, $L_{X}$, of a sample of elliptical
galaxies and their associated SMBH masses to infer tentative evidence for a
SMBH occupation fraction less than unity in galaxies with stellar
masses $M_{\star} \lesssim 10^{10}M_{\odot}$ ($M_{\bullet} \lesssim
10^{7}M_{\odot}$).  This method relies on extrapolating a power-law
fit of the $L_X-\Mbh$ distribution to low values of $L_{\rm X}$ below
the detection threshold, a potentially problematic assumption if
different physical processes control the accretion rates onto the
lowest mass SMBHs.

@@

\citet{Ho:2009a} determines the SMBH accretion rates in a sample of
early-type galaxies using X-ray observations to determine the
Bondi accretion rate and based on estimates for the mass loss rate of
evolved stars.  Both methods lead to the conclusion that the available
gas reservoir is more than sufficient to power the observed
low-luminosity AGN, assuming the standard $\sim$ 10 per cent radiative
efficiency for thin disk accretion.  Much
evidence now suggest that low-luminosity AGN result from accretion
proceeding in a radiatively inefficient mode
(\citealt{Yuan&Narayan14}), due either to the advection of
gravitationally-released energy across the SMBH horizon
(e.g.~\citealt{Narayan&Yi95}) or due to the low efficiency with which the inflowing gas ultimately reaching the vicinity of the SMBH
(e.g.~\citealt{Blandford&Begelman99}; \citealt{Li+13}).

>>>>

Understanding why most SMBHs appear to be inactive requires
characterizing their gaseous environments.  Gas near the sphere of
influence of the SMBH, hereafter the `circumnuclear medium' (CNM),
controls the mass accretion rate, $\dot{M}$.  The accretion rate in
turn sets the luminosity of the black hole and the feedback of this
energy and momentum output on larger physical scales.
The presence of dense gas in the nuclear region may lead to runaway
cooling, resulting in bursty episodes of star formation and AGN
activity (e.g.~\citealt{Ciotti&Ostriker07}; \citealt{Ciotti+10}).

Knowledge of how $\dot{M}$ depends on the SMBH mass, $\Mbh$, and other
properties of the galactic nucleus informs key questions related to
the co-evolution of SMBHs and their host galaxies with cosmic time
(e.g.~\citealt{Kormendy&Ho13}; \citealt{HeckmanBest:2014a}).  SMBH growth
in the low redshift universe is dominated by low mass black holes,
$M_{\bullet} \lesssim 10^{8}M_{\odot}$ (e.g.~\citealt{Heckman+04}), a
fact which is often attributed to the general trend of `cosmic
down-sizing' resulting from hierarchical structure growth (e.g,
\citealt{Gallo+08}).  However, the physical processes involved in
growing low mass blacks may be distinct from those operating at higher
SMBH masses and accretion rates in quasars.  A key question is
whether SMBHs grow primarily by accreting gas fed in directly from
galactic or extra-galactic scales, or whether significant growth
results also from internal, secular processes such as local stellar
mass loss in the nuclear region.

A better understanding of what mechanisms regulate accretion onto
quiescent SMBHs would have implications for a variety of observations,
such as constraints on the occupation fraction of SMBHs in low mass
galaxies.  \citet{Miller+15} use the average relationship between the
nuclear X-ray luminosities, $L_{X}$, of a sample of elliptical
galaxies and their associated SMBH masses to infer tentative evidence for a
SMBH occupation fraction less than unity in galaxies with stellar
masses $M_{\star} \lesssim 10^{10}M_{\odot}$ ($M_{\bullet} \lesssim
10^{7}M_{\odot}$).  This method relies on extrapolating a power-law
fit of the $L_X-\Mbh$ distribution to low values of $L_{\rm X}$ below
the detection threshold.  This assumption is questionable if
different physical processes control the accretion rates onto the
lowest mass SMBHs.

@@

\citet{Ho:2009a} determines the SMBH accretion rates in a sample of
early-type galaxies using X-ray observations to determine the
Bondi accretion rate and based on estimates for the mass loss rate of
evolved stars.  Both methods lead to the conclusion that the available
gas reservoir is more than sufficient to power the observed
low-luminosity AGN, assuming the standard $\sim$ 10 per cent radiative
efficiency for thin disk accretion.  Much
evidence now suggest that low-luminosity AGN result from accretion
proceeding in a radiatively inefficient mode
(\citealt{Yuan&Narayan14}), due either to the advection of
gravitationally-released energy across the SMBH horizon
(e.g.~\citealt{Narayan&Yi95}) or due to the low efficiency with which the inflowing gas ultimately reaches the SMBH
(e.g.~\citealt{Blandford&Begelman99}; \citealt{Li+13}).
